Applications of Many-Core Technologies to On-line Event Reconstruction in High Energy Physics Experiments

### 2013-11-14 papers/2013/NSS/NSS.tex
some more english updates

\subsection{Data processing in \textsc{opencl}}
We also measure the track fitting algorithm latency in an implementation  
using \textsc{opencl}. Due to limitations on the available space for storing data on the GPU, we only test the 
algorithm on up to 1000 events (\textit{i.e.} about 2 million combinations to fit). In Fig.~\ref{fig:opencl_timing} 
we show the results of algorithm latency measurements for three different modes: running the serial algorithm  
on the CPU (single-core), running the \textsc{opencl} algorithm on the CPU (multi-core), and running the \textsc{opencl} algorithm
on the AMD Radeon GPU. The \textsc{opencl} implementation of the algorithm on the CPU provides a significant speedup---about 
a factor of 5---over running the algorithm serially on the CPU. Running the same \textsc{opencl} algorithm on the GPU provides an 
even greater speedup due to the increased number of cores and parallel threads that can be run, though there is additional latency to copy data 
into and out of the GPU that makes running on the GPU take longer for small numbers of combinations to fit.
\begin{figure}[!t]
  \centering
  \includegraphics[width=0.9\linewidth]{figures/TotalTimeOpencl_1000}
  \caption{The timing of the \textsc{opencl} algorithm as a function of the
    number of track fits. We compare timing running the algorithm serially on a CPU, 
    using \textsc{opencl} on the CPU, and on an AMD GPU.}
  \label{fig:opencl_timing}
\end{figure}

@@

\section{Conclusions}
We have implemented a full version of the CDF SVT tracking algorithm
on GPUs and Intel MIC. We examined a staged approach to using
accelerator cards in a hadron collider trigger application. We have
demonstrated that in this application, significant gains can be
achieved with the `embarrassingly parallel' approach on an Intel MIC
architecture, with the smallest amount of required changes to an
existing serial code base. However, better performance is achieved
with GPUs and a more complete event-level parallelization using
\textsc{cuda}.  Additionally, we have implemented a parallelized algorithm using 
\textsc{opencl} and we have run that algorithm on a multi-core CPU and GPU, showing a boost
in performance over serial CPU computation for even small numbers of events. 
We have also updated latency studies and shown that for larger packet sizes 
(greater than 600 kB), \textsc{cuda}-aware MPI outperforms P2P. This result fully agrees with expectations 
since the P2P mechanism is optimized (and effective) to reduce the transfer latency for 
small size packets.
%We have updated latency studies and shown that for
%larger packet sizes (greater than 600~kB), \textsc{cuda}-aware MPI
%outperforms P2P, contrary to naive expectation.  
Even at large packet
size, the data transfer takes an appreciable fraction of the total
algorithm time (about 20-25\%).

>>>>

\subsection{Data processing in \textsc{opencl}}
We also measure the track fitting algorithm latency in an
implementation using \textsc{opencl}. The \textsc{opencl} tools have
the advantage of being an open standard, while the \textsc{cuda} tools
are only compatible with Nvidia GPUs. Additionally, \textsc{opencl}
can work also on CPUs and on the Xeon Phi. For our tests, due to
limitations on the available space for storing data on the GPU, we
only test the algorithm on up to 1000 events (\textit{i.e.} about 2
million combinations to fit). In Fig.~\ref{fig:opencl_timing} we show
the results of algorithm latency measurements for three different
modes: running the serial algorithm on the CPU (single-core), running
the \textsc{opencl} algorithm on the CPU (multi-core), and running the
\textsc{opencl} algorithm on the AMD Radeon GPU. The \textsc{opencl}
implementation of the algorithm on the CPU provides a significant
speedup---about a factor of 5---over running the algorithm serially on
the CPU. Running the same \textsc{opencl} algorithm on the GPU
provides an even greater speedup due to the increased number of cores
and parallel threads that can be run, though there is additional
latency to copy data into and out of the GPU that makes running on the
GPU take longer for small numbers of combinations to fit.
\begin{figure}[!t]
  \centering
  \includegraphics[width=0.9\linewidth]{figures/TotalTimeOpencl_1000}
  \caption{The timing of the \textsc{opencl} algorithm as a function of the
    number of track fits. We compare timing running the algorithm serially on a CPU, 
    using \textsc{opencl} on the CPU, and on an AMD GPU.}
  \label{fig:opencl_timing}
\end{figure}

@@

\section{Conclusions}
We have implemented a full version of the CDF SVT tracking algorithm
on GPUs and Intel MIC. We examined a staged approach to using
accelerator cards in a hadron collider trigger application and
compared additional software tool sets. We have demonstrated that in
this application, significant gains can be achieved with the
`embarrassingly parallel' approach on an Intel MIC architecture, with
the smallest amount of required changes to an existing serial code
base. Additionally, we have implemented a parallelized algorithm using
\textsc{cuda} and \textsc{opencl}. Better performance is achieved with
GPUs and a more complete event-level parallelization using these
tools.  We have run the parallelized algorithm on a multi-core CPU and
GPU, showing a boost in performance over serial CPU computation for
even small numbers of events.  We have also updated latency studies
and shown that for larger packet sizes (greater than 600 kB),
\textsc{cuda}-aware MPI outperforms P2P.  Even at large packet size,
the data transfer takes an appreciable fraction of the total algorithm
time (about 20-25\%).
